Black-swan events in animal populations

### 2016-11-17 text/ms_pnas.tex
Copyediting
Cut some text to reduce length slightly

We fit our models with Stan 2.12.1 \cite{hoffman2014, stan-manual2015, carpenter2016}, and R
3.3.1 \cite{r2015}. We began with four chains and 2,000 iterations, discarding
the first 1,000 as warm up (i.e.~4,000 final samples). If \(\hat{R}\) (the
potential scale reduction factor---a measure of chain convergence) was greater
than \(1.05\) for any parameter, or if the minimum effective sample size,
\(n_\mathrm{eff}\) (a measure of the effective number of uncorrelated samples),
for any parameter was less than \(200\), we doubled both the total iterations
and warm up period and sampled from the model again.
In the majority of cases our minimum thresholds were
greatly exceeded. We continued this procedure up to 8,000 iterations (16,000
total samples) by which all chains had sufficiently converged.
The No-U-Turn Hamiltonian Markov chain Monte Carlo sampler in Stan generally
requires fewer iterations to obtain equivalent effective sample sizes than the
Gibbs or Metropolis--Hastings algorithms \cite{stan-manual2015}.

\subsection{Alternative models and simulation testing}
We fit alternative population models to test if four key phenomena
systematically changed our conclusions. Autocorrelation has been suggested as
a reason for increased observed variability of abundance time series through
time, which could create apparent heavy tails \citep{inchausti2002};
therefore, we fit a model that included serial correlation in the residuals.
Additionally, previous work has modelled abundance or growth rates without
accounting for density dependence \citep{halley2002,segura2013}; therefore, we
fit a simpler model in which we assumed density independence.
This model is equivalent to a random walk with drift
on log abundance, and therefore does not assume
that the timeseries is stationary.
Third, observation error could bias parameter estimates \citep{knape2012} or mask our
ability to detect heavy tails \citep{ward2007}; therefore, we fit a model
where we allowed for a fixed quantity of observation error ($0.2$ standard
deviations on a log scale). Fourth, the Gompertz model assumes that
population growth rate declines linearly with log abundance. Therefore, we
also fit an alternative model, the Ricker-logistic model, which assumes that
population growth rate declines linearly with abundance (Supporting Information).
% Our models also assume that their parameters
% (e.g. parameters representing productivity and density dependence)

>>>>

We fit our models with Stan 2.12.1 \cite{hoffman2014, stan-manual2015, carpenter2016}, and R
3.3.1 \cite{r2015}. We began with four chains and 2,000 iterations, discarding
the first 1,000 as warm up. If \(\hat{R}\) (the
potential scale reduction factor---a measure of chain convergence) was greater
than \(1.05\) for any parameter, or if the minimum effective sample size,
\(n_\mathrm{eff}\) (a measure of the effective number of uncorrelated samples),
for any parameter was less than \(200\), we doubled both the total iterations
and warm up period and sampled from the model again.
In the majority of cases our minimum thresholds were
greatly exceeded. We continued this procedure up to 8,000 iterations 
by which all chains had sufficiently converged.
% The No-U-Turn Hamiltonian Markov chain Monte Carlo sampler in Stan generally
% requires fewer iterations to obtain equivalent effective sample sizes than the
% Gibbs or Metropolis--Hastings algorithms \cite{stan-manual2015}.

\subsection{Alternative models and simulation testing}
We fit alternative population models to test if four key phenomena
systematically changed our conclusions. Autocorrelation has been suggested as
a reason for increased observed variability of abundance time series through
time, which could create apparent heavy tails \citep{inchausti2002};
therefore, we fit a model that included serial correlation in the residuals.
Additionally, previous work has modelled abundance or growth rates without
accounting for density dependence \citep{halley2002,segura2013}; therefore, we
fit a simpler model in which we assumed density independence.
This model is equivalent to a random walk with drift
on log abundance, and therefore does not assume
that the time series is stationary.
Third, observation error could bias parameter estimates \citep{knape2012} or mask our
ability to detect heavy tails \citep{ward2007}; therefore, we fit a model
where we allowed for a fixed quantity of observation error ($0.2$ standard
deviations on a log scale). Fourth, the Gompertz model assumes that
population growth rate declines linearly with log abundance. Therefore, we
also fit an alternative model, the Ricker-logistic model, which assumes that
population growth rate declines linearly with abundance (Supporting Information).
% Our models also assume that their parameters
% (e.g. parameters representing productivity and density dependence)
